---
title: "Chapter 2 Lab Notes"
author: "Michael D Porter"
date: "2020-05-19"
output: html_notebook
---


The textbook and datasets are hosted online at <https://faculty.marshall.usc.edu/gareth-james/ISL/>. 
You will find that the `Auto.csv` data is at the following link <https://faculty.marshall.usc.edu/gareth-james/ISL/Auto.csv>. 

```{r}
# Load tidyverse packages
library(tidyverse)
```

Load the `Auto` data using the `read_csv()` function (from `readr` package).
```{r}
Auto = read_csv('https://faculty.marshall.usc.edu/gareth-james/ISL/Auto.csv')
```
Note: some value of `horsepower` are missing, and represented with `?`. We need to make this explictly missing (represent with an `NA`).
```{r}
Auto = read_csv('https://faculty.marshall.usc.edu/gareth-james/ISL/Auto.csv', na = "?")
```

Preview the data
```{r}
Auto
```
Summarize the data
```{r}
summary(Auto)
```

Now make boxplots. We will use the `ggplot2` package (part of tidyverse so its already loaded)
```{r}
library(ggplot2) # already loaded, but to be clear
ggplot(Auto, aes(x=factor(cylinders), y=mpg)) + 
  geom_boxplot(fill="orange", color="blue") + 
  labs(x="Number of Cylinders", y="MPG")
```

Histogram of mpg
```{r}
ggplot(Auto, aes(x=mpg)) + 
  geom_histogram(binwidth=1, center=10,
                 fill="orange", color="blue")
```






