# Chapter 2 Lab: Introduction to R

<!-- where needed I refer to the following resources -->
<!--  https://github.com/SmithCollege-SDS/tidy-islr/blob/master/lab01/Lab%201%20-%20Introduction%20to%20R.ipynb -->
<!--  https://github.com/uc-cfss/tidy-islr/blob/master/02-stat-learn.Rmd -->

## ~~Basic Commands~~ Base R

```{r base-R}
x <- c(1,3,2,5)
x
x = c(1,6,2)
x
y = c(1,4,3)
length(x)
length(y)
x+y
ls()
rm(x,y)
ls()
rm(list=ls())
?matrix
x=matrix(data=c(1,2,3,4), nrow=2, ncol=2)
x
x=matrix(c(1,2,3,4),2,2)
matrix(c(1,2,3,4),2,2,byrow=TRUE)
sqrt(x)
x^2
x=rnorm(50)
y=x+rnorm(50,mean=50,sd=.1)
cor(x,y)
set.seed(1303)
rnorm(50)
set.seed(3)
y=rnorm(100)
mean(y)
var(y)
sqrt(var(y))
sd(y)
```

## ~~Graphics~~ tidyverse

### tibble
`install.packages("tidyverse")`

```{r df}
#x=rnorm(100)
#y=rnorm(100)

library(tibble)

xy <- tibble(
  x=rnorm(100),
  y=rnorm(100)
)
```

### ggplot2

```{r scatter}
#plot(x,y)
#plot(x,y,xlab="this is the x-axis",ylab="this is the y-axis",main="Plot of X vs Y")

library(ggplot2)

scatter_plot <- ggplot(mapping=aes(x=x, y=y),
                      data=xy) +
  geom_point(color='green') +
  labs(title="Plot of X vs Y",
       x="this is the x-axis",
       y="this is the y-axis")

scatter_plot

# pdf("Figure.pdf")
# plot(x,y,col="green")
# dev.off()

ggsave(filename = "scatter_plot.pdf", plot = scatter_plot)

```

### More base R

```{r seq}
x=seq(1,10)
x
x=1:10
x
```

### More ggplot2

```{r contour}

#x=seq(-pi,pi,length=50)
#y=x
#f=outer(x,y,function(x,y)cos(y)/(1+x^2))
#contour(x,y,f)

d <- seq(-pi,pi,length=50)
x <- outer(d, d, function(x, y) x )
y <- t(x)
z <- outer(d, d, function(x, y) cos(y)/(1+x^2) )

xyz <- tibble(
  x=as.vector(x),
  y=as.vector(y),
  z1=as.vector(z)
)

contour_plot1 <- ggplot(mapping=aes(x=x, y=y, z=z1),
                        data=xyz)

contour_plot1+geom_contour()

#contour(x,y,f,nlevels=45,add=T)

contour_plot1+geom_contour()+geom_contour(bins=45)

#fa=(f-t(f))/2
#contour(x,y,fa,nlevels=15)

xyz <- xyz %>% add_column(z2=as.numeric((z-t(z))/2))
contour_plot2 <- ggplot(mapping=aes(x=x, y=y, z=z2),
                        data=xyz) 

contour_plot2+geom_contour(bins=15)

#image(x,y,fa)

contour_plot2+geom_contour_filled()
```

### plotly

`install.packages("plotly", repos="http://cran.rstudio.com")`

```{r plotly}
#persp(x,y,fa)
#persp(x,y,fa,theta=30)
#persp(x,y,fa,theta=30,phi=20)
#persp(x,y,fa,theta=30,phi=70)
#persp(x,y,fa,theta=30,phi=40)

library(plotly)

plot_ly(x=x, y=y, z=z, type='surface') %>% 
  layout(scene = list(aspectmode='cube'))

```



## ~~Indexing Data~~ More base R

```{r indexing}
A=matrix(1:16,4,4)
A
A[2,3]
A[c(1,3),c(2,4)]
A[1:3,2:4]
A[1:2,]
A[,1:2]
A[1,]
A[-c(1,3),]
A[-c(1,3),-c(1,3,4)]
dim(A)
```

## tidyverse indexing
### dplyr: filter/select

```{r dplyr}
library(dplyr)

filter(.data = xyz, x > 0)

select(.data = xyz, x, y)

```


## ~~Loading Data~~ tidyverse: readr

```{r auto1}

#Auto=read.table("Auto.data")
#fix(Auto)
#Auto=read.table("Auto.data",header=T,na.strings="?")
#fix(Auto)
#Auto=read.csv("Auto.csv",header=T,na.strings="?")
#fix(Auto)
#dim(Auto)
#Auto[1:4,]

#Load the `Auto` data using the `read_csv()` function (from `readr` package).
library(readr)
Auto = read_csv('https://faculty.marshall.usc.edu/gareth-james/ISL/Auto.csv')
Auto
Auto %>% count(horsepower)
Auto = read_csv('https://faculty.marshall.usc.edu/gareth-james/ISL/Auto.csv', 
                na = "?")
summary(Auto)

# the tidyverse can handle NA's appropriately 
# wihtout us explicitly dropping those rows

#Auto=na.omit(Auto)
#dim(Auto)
#names(Auto)

```

## ~~Additional Graphical and Numerical Summaries~~ More ggplot2

```{r auto2}

#plot(cylinders, mpg)
#plot(Auto$cylinders, Auto$mpg)
#attach(Auto)
#plot(cylinders, mpg)
#cylinders=as.factor(cylinders)
#plot(cylinders, mpg)
#plot(cylinders, mpg, col="red")
#plot(cylinders, mpg, col="red", varwidth=T)
#plot(cylinders, mpg, col="red", varwidth=T,horizontal=T)
#plot(cylinders, mpg, col="red", varwidth=T, xlab="cylinders", ylab="MPG")

ggplot(Auto, aes(x=factor(cylinders), y=mpg)) + 
  geom_boxplot(fill="orange", color="blue") + 
  labs(x="cylinders", y="MPG")

Auto <- mutate(.data=Auto,
               cylinders=as.factor(cylinders))
ggplot(data = Auto, mapping = aes(x=cylinders, y=mpg)) +
  geom_boxplot(fill="orange", color="blue") + 
  labs(x="cylinders", y="MPG")

#hist(mpg)
#hist(mpg,col=2)
#hist(mpg,col=2,breaks=15)

ggplot(Auto, aes(x=mpg)) + 
  geom_histogram(binwidth=1, 
                 center=10, 
                 fill="orange", 
                 color="blue")

```

### More tidyverse: GGally

`install.packages("GGally")`

```{r auto3}

# pairs(Auto)
# pairs(~ mpg + displacement + horsepower + weight + acceleration, Auto)

library(GGally)
ggpairs(data = select(.data = Auto, -name))

```

### More plotly (with ggplot)

```{r auto4}

#plot(horsepower,mpg)
#identify(horsepower,mpg,name)

qplot(data=Auto, x=horsepower, y=mpg, label=name)

# this converts the ggplot2 figure to an interactive plotly figure
# click on the plotly figure to enable information on hover
ggplotly()

```


### Done
```{r auto5}

# we already showed the tidyverse summary function

#summary(Auto)
#summary(mpg)