---
title: "ISLR Problem 6.9"
author: "**SOLUTIONS**"
output: html_document # uncomment and comment the following line to switch
#output: html_notebook  # notebooks may be a better option for homework
---

<!--- Below are global settings for knitr. You can override any of them by adding the changes to individual chunks --->


```{r global_options, include=FALSE}
knitr::opts_chunk$set(error=TRUE,        # Keep compiling upon error
                      collapse=FALSE,    # collapse by default
                      echo=TRUE,         # echo code by default
                      comment = "#>",    # change comment character
                      fig.width = 5.5,     # set figure width
                      fig.align = "center",# set figure position
                      out.width = "49%", # set width of displayed images
                      warning=FALSE,     # do not show R warnings
                      message=FALSE)     # do not show R messages
```

<!--- Change font size for headers --->
<style>
h1.title {
  font-size: 28px;
}
h1 {
  font-size: 22px;
}
h2 {
  font-size: 18px;
}
h3 { 
  font-size: 14px;
}

</style>



**SYS 6018 | Summer 2020 | University of Virginia **

Solutions to the homework problems in [Introduction to Statistical Learning (ISLR)](https://faculty.marshall.usc.edu/gareth-james/ISL/).

*******************************************

Solutions will be shown using Base R (following the textbook) and the [Tidyverse](https://www.tidyverse.org/learn/) dialect of R. Either (or a mix of both) is acceptable, but I encourage tidyverse for those who can take the extra time to deviate from the textbook. It is well worth the investment.

To use tidyverse solutions, we will need to load the following packages:
```{r, warning=FALSE}
library(tidyverse)  # ggplot2, dplyr, etc.
library(broom)      # to get tidy model output
library(knitr)      # for the kable() function for tables
library(kableExtra) # to make fancier html tables
library(glmnet)     # for ridge and lasso regression
library(pls)        # for pcr and pls
```

In this exercise, we will predict the number of applications received using the other variables in the `College` data set.

Load the `College` data.  
```{r load-data}
#-- Load data. Convert to tibble for better printing. Add rownames.
College = ISLR::College %>% 
  as_tibble(rownames = "College")

#-- Load csv directly from web:
# College = read_csv("https://faculty.marshall.usc.edu/gareth-james/ISL/College.csv") %>%
#   rename(College=X1)
```


## (a).  Split the data set into a training set and a test set. {.tabset}

There are `r nrow(College)` observations which isn't alot considering we have `r ncol(College)` predictors. That is `r {nrow(College)/ncol(College)} %>% round(1)` observations per parameter (if one parameter per predictor). I'll reserve 50  (`r 100*round(50/nrow(College), 3)`%) observations for test and use the rest for training. 

### Tidyverse
```{r}
#-- Settings
n = nrow(College)
n.test = 50
n.train = n - n.test

#-- select training data
set.seed(61)
train = c(rep(TRUE, n.train), rep(FALSE, n.test)) %>%
  sample()

```


### Base R

```{r}
#-- Settings
n = nrow(College)
n.test = 50
n.train = n - n.test

#-- select training data
set.seed(61)
train = sample(c(rep(TRUE, n.train), rep(FALSE, n.test)))
```



## (b). Linear Model 

Fit a linear model using least squares on the training set, and report the test error obtained.

Note: There is an annoying feature with `lm()` and `predict.lm()`. An unused predictor, like `College` is still carried along in the model and converted to a factor. Thus the factor levels won't match up when trying to predict and you will get a strange error message. Thus I am going to create a data frame called data that removes the `College` variable. 

```{r}
data = College %>% select(-College)
```

Now everything works as expected:
```{r}
#-- Formula
fmla = formula(Apps ~ .)  # everything used as predictors

#-- Linear Model (unpenalized)
fit.lm = lm(fmla,
            data=data[train, ])

#-- Prediction on test set
yhat.lm = predict(fit.lm, newdata=data[!train,])
```

And test error (I'll use RMSE since its not specified what to use)
```{r}
rmse <- function(yhat, y) {
  sqrt(mean((y - yhat)^2))
}
```

```{r}
#-- test error
rmse(yhat.lm, data$Apps[!train])
```


## (c). Ridge Regression 

Fit a ridge regression model on the training set, with $\lambda$ chosen by cross-validation. Report the test error obtained.

Note: the `glmnet` package requires matrix input instead of formulas and data frames. Thus, we can use `model.matrix()` to get our matrices. 

```{r}
#-- Packages
library(glmnet)

#-- Set up design matrix
X = model.matrix(fmla, data=data)[,-1] # remove Intercept
Y = data$Apps
```


To ensure all folds are consistent over all models, the folds need to be set manually. Here I'll use 10-fold cross-validation. 
```{r}
#-- Set folds so consistent over all runs
set.seed(62)     # set seed for reproducibility
K = 10           # number of folds
folds = rep(1:K, length=sum(train)) # make folds
```

Now we can estimate and predict
```{r}
#-- Ridge Regression
fit.ridge = cv.glmnet(x=X[train,], y=Y[train],
                      alpha=0,    # ridge penalty
                      foldid=folds)

print(fit.ridge)

#-- predict on test set
yhat.ridge = predict(fit.ridge, 
                            s="lambda.min",
                            newx=X[!train,]) 
```

```{r}
#-- test error
rmse(yhat.ridge, Y[!train])
```



## (d). Lasso Regression 

Fit a lasso model on the training set, with $\lambda$ chosen by cross-validation. Report the test error obtained, along with the number of non-zero coeﬃcient estimates.

Estimate and predict
```{r}
#-- Ridge Regression
fit.lasso = cv.glmnet(x=X[train,], y=Y[train],
                      alpha=1,    # lasso penalty
                      foldid=folds)

print(fit.lasso)

#-- predict on test set
yhat.lasso = predict(fit.lasso, 
                            s="lambda.min",
                            newx=X[!train,]) 
```

```{r}
#-- test error
rmse(yhat.lasso, Y[!train])
```


## (e). Principal Component Regression 

Fit a PCR model on the training set, with M chosen by cross-validation. Report the test error obtained, along with the value of M selected by cross-validation.


The `pls` package doesn't appear to allow specification of the folds so I'll manually run CV for this model
```{r}
#-- load package
library(pls)

#-- Cross-validation
ind.train = which(train)   # index of training observations
  
sse = matrix(NA, K, 17)
for(k in unique(folds)) {
  ind.fit = ind.train[folds != k]
  ind.val = ind.train[folds == k]
  pcr.fit = pcr(fmla, data=data[ind.fit,],
              scale=TRUE)
  yhat = predict(pcr.fit, data[ind.val,])[,1,]
  y.true = data$Apps[ind.val]
  sse[k,] = apply(yhat, 2, function(x) sum( (y.true-x)^2 ))
}

tibble(ncomp = 1:17, sse = colSums(sse)) %>% 
  arrange(sse)

#-- Or ignore the folds and let the model choose
pcr(fmla, data=data[train, ], validation="CV") %>%
  summary()

```

It appears the full model (same as least-squares) is best! Thus same error as least squares.  


## (f). Partial Least Squares 

Fit a PLS model on the training set, with M chosen by cross- validation. Report the test error obtained, along with the value of M selected by cross-validation.

```{r}
sse = matrix(NA, K, 17)
for(k in unique(folds)) {
  ind.fit = ind.train[folds != k]
  ind.val = ind.train[folds == k]
  plsr.fit = plsr(fmla, data=data[ind.fit,],
              scale=TRUE)
  yhat = predict(plsr.fit, data[ind.val,])[,1,]
  y.true = data$Apps[ind.val]
  sse[k,] = apply(yhat, 2, function(x) sum( (y.true-x)^2 ))
}

tibble(ncomp = 1:17, sse = colSums(sse)) %>% 
  arrange(sse)
```

This happens to select 9 components. The error is
```{r}
plsr.fit = plsr(fmla, data=data[train, ])
yhat.plsr = predict(plsr.fit, data[!train, ], ncomp=9)[,1,1]
rmse(yhat.plsr, Y[!train])
```


## (g). Results 

Comment on the results obtained. How accurately can we predict the number of college applications received? Is there much difference among the test errors resulting from these five approaches?

```{r}
yhat = tibble(lm = yhat.lm, 
              ridge=yhat.ridge, 
              lasso=yhat.lasso, 
              pcr = yhat.lm, 
              plsr = yhat.plsr)
apply(yhat, 2, rmse, y=Y[!train])
```


The winner goes to the full linear model. Using this particular train/test split the penalized/reduced models can't predict as well as the full model. Perhaps there would be a slightly different result if we used a different split. 





