---
title: "Mod_1_Live"
author: "Schwartz"
date: "8/30/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE,
                      fig.align = "center",
                      out.width = "80%")

```

# HIT RECORD
# HIT RECORD
# HIT RECORD

## Time Demands

## Module Section Prefaces

## Project Collaboration Rules

## Demos

# HIT RECORD
# HIT RECORD
# HIT RECORD


## College Data
```{r college, fig.width = 5}
library("tidyverse")
library("modelr")
url <- "http://faculty.marshall.usc.edu/gareth-james/ISL/College.csv"
college <- read_csv(url, col_types = cols()) %>% 
  rename(college=X1) %>%
  mutate(Accept.Rate = Accept/Apps)

set.seed(10)
college.train <- college %>% 
  filter(Private!="Yes") %>% dplyr::select(-Private) %>% 
  sample_n(100, replace=FALSE)
college.test <- college %>% 
  filter(Private=="Yes") %>% dplyr::select(-Private) %>%
  sample_n(100, replace=FALSE)

# grepl(pattern, x)
college.train %>% filter(str_detect(college, 'Virginia') | 
                         str_detect(college, 'George') |
                         str_detect(college, 'James'))
```

## Model Complexity
```{r complexity}
p='7'
```

## Training Data Set I
```{r ds1, fig.width = 3}
my_plot <- college.train %>% 
  ggplot(mapping=aes(x=Outstate, y=Accept.Rate)) +
  geom_point(color='forestgreen') 
my_plot
```

## Model Fit I (Training Data Set I)
```{r mf1, fig.width = 3}
my_plot + stat_smooth(method="lm", formula=paste0('y~poly(x,', p, ')'), 
                      se=FALSE, color='forestgreen')
```

## Training Data Set II
```{r ds2, fig.width = 4}
my_plot + stat_smooth(method="lm", formula=paste0('y~poly(x,', p, ')'), 
                      se=FALSE, color='forestgreen') + 
  geom_point(data=college.test, color='purple', size=2)
```


## Model Fit II (Training Data Set II)
```{r mf2, fig.width = 4}
my_plot + stat_smooth(method="lm", formula=paste0('y~poly(x,', p, ')'), 
                      se=FALSE, color='forestgreen') + 
  geom_point(data=college.test, color='purple') + 
  stat_smooth(data=college.test, method="lm", 
              formula=paste0('y~poly(x,', p, ')'), se=FALSE, color='purple')
```


## Confidence Intervals
```{r ci, fig.width = 4}
my_plot <- 
my_plot + stat_smooth(method="lm", formula=paste0('y~poly(x,', p, ')'), 
                      se=TRUE, color='forestgreen') + 
  geom_point(data=college.test, color='purple') + 
  stat_smooth(data=college.test, method="lm", 
              formula=paste0('y~poly(x,', p, ')'), se=TRUE, color='purple')
my_plot
```

## Prediction Intervals I: (Model Fit I)
```{r pi1, fig.width = 4}
mod1 <- lm(Accept.Rate~poly(Outstate, strtoi(p)), data=college.train) 
mod1.support <- college.train %>% dplyr::select(Outstate) %>% seq_range(n=20)
mod1.prediction_interval <- tibble(Outstate=mod1.support)
mod1.prediction_interval <- 
predict(mod1, newdata=mod1.prediction_interval, interval="predict") %>%
  as_tibble() %>% add_column(mod1.prediction_interval) %>%
  rename(Accept.Rate=fit)
my_plot + geom_ribbon(data=mod1.prediction_interval, 
                      aes(x=Outstate, ymin=lwr, ymax=upr), 
                      fill = "black", alpha = 0.2)
```  

## Prediction Intervals II: (Model Fit II)
```{r pi2, fig.width = 4}
mod2 <- lm(Accept.Rate~poly(Outstate, strtoi(p)), data=college.test) 
mod2.support <- college.test %>% dplyr::select(Outstate) %>% seq_range(n=20)
mod2.prediction_interval <- tibble(Outstate=mod2.support)
mod2.prediction_interval <- 
predict(mod2, newdata=mod2.prediction_interval, interval="predict") %>%
  as_tibble() %>% add_column(mod2.prediction_interval) %>%
  rename(Accept.Rate=fit)
my_plot + geom_ribbon(data=mod2.prediction_interval, 
                      aes(x=Outstate, ymin=lwr, ymax=upr), 
                      fill = "black", alpha = 0.2)
```

## Plot Clean-Up
```{r clean-up, fig.width = 4}
my_plot$layers[[1]] <- NULL
my_plot$layers[[2]] <- NULL
```

## Resample Training Data Sets I & II
```{r college, fig.width = 4}
resample <- function(college){
  college.train.2 <- college %>% 
    filter(Private!="Yes") %>% dplyr::select(-Private) %>% 
    sample_n(100, replace=FALSE)
  college.test.2 <- college %>% 
    filter(Private=="Yes") %>% dplyr::select(-Private) %>%
    sample_n(100, replace=FALSE)
  list(college.train.2, college.test.2)
}
```

# new model fits
```{r re-fits, fig.width = 4}
parallel_universe <- resample(college)

mod1.2 <- lm(Accept.Rate~poly(Outstate, strtoi(p)), data=parallel_universe[[1]]) 
mod1.2.support <- college.train %>% dplyr::select(Outstate) %>% seq_range(n=20)
mod1.2.fit <- mod1.2 %>% augment(newdata=tibble(Outstate=mod1.2.support))

mod2.2 <- lm(Accept.Rate~poly(Outstate, strtoi(p)), data=parallel_universe[[2]])
mod2.2.support <- college.test %>% dplyr::select(Outstate) %>% seq_range(n=20)
mod2.2.fit <-  mod2.2 %>% augment(newdata=tibble(Outstate=mod2.2.support))

my_plot + geom_line(data=mod1.2.fit, mapping=aes(x=Outstate, y=.fitted),
                    size=1, linetype='dotdash') +
          geom_line(data=mod2.2.fit, mapping=aes(x=Outstate, y=.fitted),
                    size=1, linetype='dotdash') 
```



## *My* Questions
https://pollev.com/scottschwart658

## *Your* Questions

