# Chapter 7 Lab: Non-linear Modeling

```{r global_options, include=FALSE}
knitr::opts_chunk$set(error=TRUE,        # Keep compiling upon error
                      collapse=FALSE,    # collapse by default
                      echo=TRUE,         # echo code by default
                      comment = "#>",    # change comment character
                      fig.width = 5.5,     # set figure width
                      fig.align = "center",# set figure position
                      out.width = "80%", # set width of displayed images
                      warning=FALSE,     # do not show R warnings
                      message=FALSE)     # do not show R messages
```


```{r setup}
#library(ISLR)
#attach(Wage)
library(tidyverse)
```

## Polynomial Regression and Step Functions

I have not made a final hold out test data set... 
If we're doing classical statistics then we're relying on uncertainty 
characterization as opposed to model tuning approaches.  

```{r data, fig.width=8}
#fit=lm(wage~poly(age,4),data=Wage)
#coef(summary(fit))
fit <- lm(wage~poly(age,4), data=ISLR::Wage)
fit %>% broom::tidy()

#fit2=lm(wage~poly(age,4,raw=T),data=Wage)
#coef(summary(fit2))
fit2 <- lm(wage~poly(age,4,raw=T), data=ISLR::Wage)
fit2 %>% broom::tidy()

#fit2a=lm(wage~age+I(age^2)+I(age^3)+I(age^4),data=Wage)
#coef(fit2a)
fit2a <- lm(wage~age+I(age^2)+I(age^3)+I(age^4), data=ISLR::Wage)
fit2a %>% broom::tidy()

#fit2b=lm(wage~cbind(age,age^2,age^3,age^4),data=Wage)
fit2b <- lm(wage~cbind(age,age^2,age^3,age^4), data=ISLR::Wage)
fit2b %>% broom::tidy()

#agelims=range(age)
#age.grid=seq(from=agelims[1],to=agelims[2])

age_poly <- poly(ISLR::Wage$age, degree = 4)
X <- age_poly %>% as_tibble() %>%
  set_names(paste0("age_poly",as.character(1:4)))
wage <- ISLR::Wage %>% as_tibble()
wage[,names(X)] <- X

fit <- lm(wage~age_poly1+age_poly2+age_poly3+age_poly4, data=wage)
fit %>% broom::tidy()

age.grid <- wage %>% dplyr::select(age) %>% modelr::seq_range(n=30)
age_poly.grid <- predict(age_poly, newdata=age.grid) %>% as_tibble() %>%
  set_names(paste0("age_poly",as.character(1:4))) %>% add_column(age=age.grid)

#preds=predict(fit,newdata=list(age=age.grid),se=TRUE)
#se.bands=cbind(preds$fit+2*preds$se.fit,preds$fit-2*preds$se.fit)

#install.packages("tidypredict")
library(tidypredict)
prediction_interval <- age_poly.grid %>% 
  mutate(pred=!!tidypredict_fit(fit),
         pm_int=!!tidypredict_interval(fit, interval=0.95)) %>%
  mutate(lower=pred-pm_int, upper=pred+pm_int)

#par(mfrow=c(1,2),mar=c(4.5,4.5,1,1),oma=c(0,0,4,0))
#plot(age,wage,xlim=agelims,cex=.5,col="darkgrey")
#title("Degree-4 Polynomial",outer=T)
#lines(age.grid,preds$fit,lwd=2,col="blue")
#matlines(age.grid,se.bands,lwd=1,col="blue",lty=3)

# https://stackoverflow.com/questions/7714677/scatterplot-with-too-many-points
ggplot() + geom_hex(wage, mapping=aes(x=age, y=wage)) + 
  geom_jitter(wage, mapping=aes(x=age, y=wage),
              shape='.', size=.1, color='white', height=2) +
  geom_ribbon(data=prediction_interval, fill = "black", alpha = 0.2, 
              mapping=aes(x=age, ymin=lower, ymax=upper)) + 
  stat_smooth(wage, mapping=aes(x=age, y=wage),
              method='lm', formula='y~poly(x,4)', se=TRUE, alpha=1) +
  geom_line(prediction_interval, mapping=aes(x=age, y=pred)) -> fit.plot
              
# preds2=predict(fit2,newdata=list(age=age.grid),se=TRUE)
preds2 <- predict(fit2, newdata=list(age=age.grid), se=TRUE)
#max(abs(preds$fit-preds2$fit))
max(abs(preds2$fit-prediction_interval$pred))

fit.1 <- lm(wage~age, data=ISLR::Wage)
fit.2 <- lm(wage~poly(age,2), data=ISLR::Wage)
fit.3 <- lm(wage~poly(age,3), data=ISLR::Wage)
fit.4 <- lm(wage~poly(age,4), data=ISLR::Wage)
fit.5 <- lm(wage~poly(age,5), data=ISLR::Wage)
anova(fit.1,fit.2,fit.3,fit.4,fit.5)

#coef(summary(fit.5))
#(-11.983)^2
fit.5 %>% tidy()
fit.5 %>% tidy() %>% slice(3) %>% select(statistic) %>% 
  mutate(F_score=statistic^2) %>% select(F_score) %>% pull()

fit.1 <- lm(wage~education+age, data=ISLR::Wage)
fit.2 <- lm(wage~education+poly(age,2), data=ISLR::Wage)
fit.3 <- lm(wage~education+poly(age,3), data=ISLR::Wage)
anova(fit.1,fit.2,fit.3)

#fit=glm(I(wage>250)~poly(age,4),data=Wage,family=binomial)
wage %>% mutate(rich = wage>250) -> wage
fit_glm <- glm(rich~age_poly1+age_poly2+age_poly3+age_poly4, 
               data=wage, family=binomial)
preds <- fit_glm %>% predict(newdata=age_poly.grid, se=TRUE)
pfit <- exp(preds$fit)/(1+exp(preds$fit))
se.bands.logit <- cbind(preds$fit+2*preds$se.fit, preds$fit-2*preds$se.fit)
se.bands <- exp(se.bands.logit)/(1+exp(se.bands.logit))

#preds=predict(fit,newdata=list(age=age.grid),type="response",se=T)
#plot(age,I(wage>250),xlim=agelims,type="n",ylim=c(0,.2))
#points(jitter(age), I((wage>250)/5),cex=.5,pch="|",col="darkgrey")
#lines(age.grid,pfit,lwd=2, col="blue")
#matlines(age.grid,se.bands,lwd=1,col="blue",lty=3)
fit_glm.pred <- se.bands %>% as_tibble() %>% set_names(c("upper","lower")) %>%
  add_column(pfit=pfit)
fit_glm.pred[,names(age_poly.grid)] <- age_poly.grid

ggplot() + geom_line(fit_glm.pred, mapping=aes(x=age, y=pfit)) +
  geom_ribbon(fit_glm.pred, mapping=aes(x=age, ymin=lower, ymax=upper),
              alpha=0.5, fill='blue') +
  geom_jitter(wage%>%mutate(rich=(as.numeric(rich)-0.5)/5), 
              shape='|', mapping=aes(x=age, rich), height=.025) -> fit_glm.plot
  
gridExtra::grid.arrange(grobs=list(fit.plot,fit_glm.plot), ncol=2)

#table(cut(age,4))
#fit=lm(wage~cut(age,4),data=Wage)
#coef(summary(fit))
table(cut(ISLR::Wage$age,4))
lm(wage~cut(age,4), data=ISLR::Wage) %>% tidy()
```



## Splines

A (degree 2) truncated power basis function gives us

|derivative|function            | | regime     |
|--------|----------------------|-|------------|
|$f(x)$  |$x+x^2$               |:|$x<\epsilon$|
|$f(x)$  |$x+x^2+(x-\epsilon)^2$|:|$x>\epsilon$|
|$f'(x)$ |$2x$                  |:|$x<\epsilon$|
|$f'(x)$ |$2x+(2x-2\epsilon)$   |:|$x>\epsilon$|
|$f''(x)$|$2$                   |:|$x<\epsilon$|
|$f''(x)$|$4$                   |:|$x>\epsilon$|

and you can see that the discontinuity only appears at $\epsilon$ in the 
second derivative of the function.

```{r bsplines}
# B-Splines, however, are just a different way to construct a set of basis 
# functions around your selected knots that provide the same degree of 
# continuity in your fitted response curve:
# https://stats.stackexchange.com/questions/168351/truncated-power-basis-function-and-continuity-in-b-splines

#library(splines)
#fit=lm(wage~bs(age,knots=c(25,40,60)),data=Wage)

B_Spline <- splines::bs(wage$age, knots=c(25,40,60)) 
B_Spline_X <- B_Spline %>% as_tibble() %>%
  set_names(paste0('degree3Bspline',as.character(1:dim(B_Spline)[2])))
wage[,names(B_Spline_X)] <- B_Spline_X

linear_form <- formula(paste0("wage~", paste0(names(B_Spline_X), collapse="+")))
fit_bspline <- lm(linear_form , data=wage)

#pred=predict(fit,newdata=list(age=age.grid),se=T)

age_bspline.grid <- predict(B_Spline, newx=age.grid) %>% 
  as_tibble() %>% set_names(names(B_Spline_X)) %>% add_column(age=age.grid)
#predict(fit_bspline, newdata=age_bspline.grid, se=TRUE)
prediction_interval_bspline <- age_bspline.grid %>% 
  mutate(pred=!!tidypredict_fit(fit_bspline),
         pm_int=!!tidypredict_interval(fit_bspline, interval=0.95)) %>%
  mutate(lower=pred-pm_int, upper=pred+pm_int)

#plot(age,wage,col="gray")
#lines(age.grid,pred$fit,lwd=2)
#lines(age.grid,pred$fit+2*pred$se,lty="dashed")
#lines(age.grid,pred$fit-2*pred$se,lty="dashed")

# https://stackoverflow.com/questions/7714677/scatterplot-with-too-many-points
ggplot() + 
  stat_density_2d(wage, mapping=aes(x=age, y=wage, fill=stat(density)), 
                  geom='raster', contour=FALSE) +  
  scale_fill_viridis_c() +
  geom_jitter(wage, mapping=aes(x=age, y=wage), 
              shape='.', color='white', height=2, width=1.5) +
  geom_ribbon(prediction_interval_bspline, alpha=0.2, fill='white',
              mapping=aes(x=age, ymin=lower, ymax=upper)) +
    stat_smooth(wage, mapping=aes(x=age, y=wage),
              method='loess', formula='y~x', se=TRUE, alpha=1) +
  geom_line(prediction_interval_bspline, mapping=aes(x=age, y=pred)) -> 
  splines.plot
```


Natural splines force the fitted response surface to be first order (linear)
at the extremes of any knots:

$$x\beta_1 + (x-k_1)^3\beta_2 1_{[k_1,k_2)}(x) + (x-k_2)^3\beta_3 1_{[k_2,k_3)}(x) + 
(x-k_3)\beta_3 1_{[k_3,\infty)}(x)$$

```{r nslines}
#dim(bs(age,knots=c(25,40,60)))
#dim(bs(age,df=6))
#attr(bs(age,df=6),"knots")
dim(splines::bs(wage$age, knots=c(25,40,60)))
dim(splines::bs(wage$age, df=6))
attr(splines::bs(wage$age, df=6), "knots")

#fit2=lm(wage~ns(age,df=4),data=Wage)

N_Spline <- splines::ns(wage$age, df=4)
N_Spline_X <- N_Spline %>% as_tibble() %>%
  set_names(paste0('naturalCubicBasisFunction',as.character(1:dim(N_Spline)[2])))
wage[,names(N_Spline_X)] <-N_Spline_X

linear_form <- formula(paste0("wage~", paste0(names(N_Spline_X), collapse="+")))
fit_nspline <- lm(linear_form , data=wage)

#pred2=predict(fit2,newdata=list(age=age.grid),se=T)

age_nspline.grid <- predict(N_Spline, newx=age.grid) %>% 
  as_tibble() %>% set_names(names(N_Spline_X)) %>% add_column(age=age.grid)
#predict(fit_bspline, newdata=age_bspline.grid, se=TRUE)
prediction_interval_nspline <- age_nspline.grid %>% 
  mutate(pred=!!tidypredict_fit(fit_nspline),
         pm_int=!!tidypredict_interval(fit_nspline, interval=0.95)) %>%
  mutate(lower=pred-pm_int, upper=pred+pm_int)

#lines(age.grid, pred2$fit,col="red",lwd=2)
splines.plot + geom_line(prediction_interval_nspline, 
                         mapping=aes(x=age, y=pred), color='red')
```     

```{r smoothing}
#plot(age,wage,xlim=agelims,cex=.5,col="darkgrey")
#title("Smoothing Spline")
#fit=smooth.spline(age,wage,df=16)
#fit2=smooth.spline(age,wage,cv=TRUE)
#fit2$df
#lines(fit,col="red",lwd=2)
#lines(fit2,col="blue",lwd=2)
#legend("topright",legend=c("16 DF","6.8 DF"),col=c("red","blue"),lty=1,lwd=2,cex=.8)

#plot(age,wage,xlim=agelims,cex=.5,col="darkgrey")
#title("Local Regression")
#fit=loess(wage~age,span=.2,data=Wage)
#fit2=loess(wage~age,span=.5,data=Wage)
#lines(age.grid,predict(fit,data.frame(age=age.grid)),col="red",lwd=2)
#lines(age.grid,predict(fit2,data.frame(age=age.grid)),col="blue",lwd=2)
#legend("topright",legend=c("Span=0.2","Span=0.5"),col=c("red","blue"),lty=1,lwd=2,cex=.8)

# https://stackoverflow.com/questions/7714677/scatterplot-with-too-many-points
ggplot() + geom_density_2d(wage, mapping=aes(x=age, y=wage, color='data')) +
  ggtitle("Smoothing Splines and Loess") +
  geom_path(as_tibble(predict(smooth.spline(wage$age, wage$wage, df=16))),
            mapping=aes(x=x,y=y, color='Smoothing Spline: 16 df'), size=1) +
  geom_path(as_tibble(predict(smooth.spline(wage$age, wage$wage, cv=TRUE))),
            mapping=aes(x=x,y=y,color='Smoothing Spline: CV df'), size=1) +
  geom_smooth(wage, mapping=aes(x=age, y=wage, color='Loess (span=0.2)'),
              method='loess', formula='y~x', span=0.2, se=FALSE) +
  geom_smooth(wage, mapping=aes(x=age, y=wage, color='Loess (span=0.5)'),
              method='loess', formula='y~x', span=0.5, se=FALSE) 

```


# GAMs

```{r gams, fig.height=2}

gam1 <- lm(wage~splines::ns(year,4)+splines::ns(age,5)+education, data=wage)
library(gam)

gam.m3 <- gam(wage~s(year,4)+s(age,5)+education, data=wage)
par(mfrow=c(1,3))
plot(gam.m3, se=TRUE, col="blue")

plot.Gam(gam1, se=TRUE, col="red")

gam.m1 <- gam(wage~s(age,5)+education, data=wage)
gam.m2 <- gam(wage~year+s(age,5)+education, data=wage)
anova(gam.m1,gam.m2,gam.m3, test="F")

summary(gam.m3)

preds <- predict(gam.m2, newdata=wage)

gam.lo <- gam(wage~s(year,df=4)+lo(age,span=0.7)+education, data=wage)
plot.Gam(gam.lo, se=TRUE, col="green")
```

```{r gams_2}
gam.lo.i <- gam(wage~lo(year,age,span=0.5)+education, data=wage)
library(akima)
#par(mfrow=c(1,2))
plot(gam.lo.i, theta=75, phi=-30)
```


```{r gams_3, fig.height=2}
gam.lr<-gam(I(wage>250)~year+s(age,df=5)+education, family=binomial, data=wage)
par(mfrow=c(1,3))
plot(gam.lr, se=TRUE, col="green")

table(wage$education, I(wage$wage>250))

gam.lr.s <- gam(I(wage>250)~year+s(age,df=5)+education, 
                family=binomial, data=wage, subset=(education!="1. < HS Grad"))
plot(gam.lr.s, se=TRUE, col="green")

```










